---
title: "R Notebook"
output: html_notebook
---

#Research Qustions:

1.	What does descriptive statistics of the selected variables look like?(With certain graphs to describe)
2.	Which department has the highest attrition rate?
3.	What’s the relationship between HourlyRate and MonthlyIncome ?
4.	Could we predict on the JobSatisfaction based on EnvironmentSatisfaction, MonthlyIncome, WorkLifeBalance and their interaction on it?
5.	Can we build up a decision tree that helps predict Attrition in IBM?


```{r}
library(tidyr)
library(dplyr)
library(psych)
library(ggplot2)
library(rpart)
library(rpart.plot)

citation(package = "tidyr")

```




```{r}
df<-read.csv("BMHR.csv",stringsAsFactors = FALSE, header = TRUE,encoding = "utf-8")
colname<-names(df)
colname[1]<-"Age"
names(df)<-colname
describe(df)# the data is quite complete without any missing data
str(df)

sum(duplicated(df))
```
```{r}
# descriptive graph
# overall age and gender distribution
ggplot(df,aes(x=Age,fill=Gender))+
  geom_histogram(alpha=0.6,binwidth = 1.4)+
  facet_grid(~Gender)

# diff(range(df$Age))/30=1.4 to solve the binned problem

# Overall IBM has more males than females,of which a large part age from 25 to 35

# Department and hourlyrate and daily rate by dept
ggplot(df,aes(x=DailyRate,y=HourlyRate,col=Department))+
  geom_point()+
  scale_fill_brewer(palette = "Set1")
# It seems no certain pattern relating hourly rate and daily rate

#hourly rate and performance by job satisfaction
quantile(df$HourlyRate)
#30.00  48.00  66.00  83.75 100.00 setting threshould to match certain hourlyrate labels for 

df<-df%>%
  mutate(HourlyRateCat=case_when(
    HourlyRate <= 48 ~ "Bottom",
    HourlyRate <= 66 & HourlyRate > 48 ~ "Medium low",
    HourlyRate <= 83.75 & HourlyRate > 66 ~ "Medium high",
    HourlyRate <= 100 & HourlyRate > 83.75 ~ "Top"
))

ggplot(df,aes(x=HourlyRateCat,fill=OverTime))+
  geom_bar(position="fill",width=0.5)
#It seems that hourlyrate has too much to do with whether overtime or not

#WorkLifeBalance+OverTime+YeasAtCompany+MonthlyRate
ggplot(df,aes(x=YearsAtCompany,y=MonthlyRate,col=OverTime,size=WorkLifeBalance))+ 
  geom_point(position="jitter",alpha=0.4)+
  geom_smooth()

#How people in our company are satisfied about their job by Job Role
library(RColorBrewer)
blues <- brewer.pal(9, "Blues")
ggplot(df,aes(x=JobRole,fill=factor(JobSatisfaction)))+
  geom_bar(width=0.5)+
  scale_fill_manual(values=blues)

#How people in our company are satisfied about their job by Department
greens <- brewer.pal(4, "Greens")

ggplot(df,aes(x=Department,fill=factor(JobSatisfaction)))+
  geom_bar(position="dodge",aplha=0.8)+
  scale_fill_manual(values=greens)

#Performance
ggplot(df,aes(x=Department,fill=OverTime))+
  geom_bar(width = 0.3) # The overtime rate has been controlled in a positive way.

#Attrition

  
```



```{r}
library(ggplot2)
ggplot(df,aes(x=Age,fill=Gender))+
  geom_bar()

#dummy coding: if attrition is "Yes", coded as 0, whereas "No" as 1
df$Attrition<-ifelse(df$Attrition=="No",1,0)

#Attrition Rate
1-mean(df$Attrition) # 16.1% attrition seems okay, 10% is optimal.

#Check whether it is a good attrition or bad
df_attrition<-filter(df,df$Attrition==0)

t<-table(df_attrition$PerformanceRating,df_attrition$Department)
class(t)

ggplot(t,aes())+
  coord_polar("y", start=0)

prop.table(x=t,margin = 1)


#df%>%
#  group_by(Department)%>%
#  summarise(Attrition_rate=sum(df$Attrition))


```



```{r}
nrow(df)*0.75
sample_size<-sample(nrow(df),nrow(df)*0.75)

#set up training and 
training<-df[sample_size,]
testing<-df[-sample_size,]
pre_prune<-rpart.control(maxdepth = 30,cp=0)
model<-rpart(Attrition~.,training,method="class",control = pre_prune)



#post_pruning to get a more accurate model
plotcp(model)
post_prune<-prune(model,cp=0.013)

#draw a plot of 
rpart.plot(post_prune)
printcp(post_prune)


# test the model accuracy
testing$pre<-predict(post_prune,testing,"class")
mean(testing$pre==testing$Attrition)




```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
