---
title: "R Notebook"
output: html_notebook
---

#Research Qustions:

1.	What does descriptive statistics of the selected variables look like?(With certain graphs to describe)
2.	Which department has the highest attrition rate?
3.	What’s the relationship between HourlyRate and MonthlyIncome ?
4.	Could we predict on the JobSatisfaction based on EnvironmentSatisfaction, MonthlyIncome, WorkLifeBalance and their interaction on it?
5.	Can we build up a decision tree that helps predict Attrition in IBM?


```{r}
library(tidyr)
library(dplyr)
library(psych)
library(ggplot2)
library(rpart)
library(rpart.plot)

citation(package = "tidyr")

```

```{r}
df<-read.csv("BMHR.csv",stringsAsFactors = FALSE, header = TRUE,encoding = "utf-8")
colname<-names(df)
colname[1]<-"Age"
names(df)<-colname
describe(df)# the data is quite complete without any missing data
str(df)

sum(duplicated(df))
```
```{r}
# descriptive graph
# overall age and gender distribution
ggplot(df,aes(x=Age,fill=Gender))+
  geom_histogram(alpha=0.6,binwidth = 1.4)+
  facet_grid(~Gender)

# diff(range(df$Age))/30=1.4 to solve the binned problem

# Overall IBM has more males than females,of which a large part age from 25 to 35

# Department and hourlyrate and daily rate by dept
ggplot(df,aes(x=DailyRate,y=HourlyRate,col=Department))+
  geom_point()+
  scale_fill_brewer(palette = "Set1")
# It seems no certain pattern relating hourly rate and daily rate

#hourly rate and performance by job satisfaction
quantile(df$HourlyRate)
#30.00  48.00  66.00  83.75 100.00 setting threshould to match certain hourlyrate labels for 

df<-df%>%
  mutate(HourlyRateCat=case_when(
    HourlyRate <= 48 ~ "Bottom",
    HourlyRate <= 66 & HourlyRate > 48 ~ "Medium low",
    HourlyRate <= 83.75 & HourlyRate > 66 ~ "Medium high",
    HourlyRate <= 100 & HourlyRate > 83.75 ~ "Top"
))

ggplot(df,aes(x=HourlyRateCat,fill=OverTime))+
  geom_bar(position="fill",width=0.5)
#It seems that hourlyrate has too much to do with whether overtime or not

#WorkLifeBalance+OverTime+YeasAtCompany+MonthlyRate
ggplot(df,aes(x=YearsAtCompany,y=MonthlyRate,col=OverTime,size=WorkLifeBalance))+ 
  geom_point(position="jitter",alpha=0.4)+
  geom_smooth()

#How people in our company are satisfied about their job by Job Role
library(RColorBrewer)
blues <- brewer.pal(9, "Blues")
ggplot(df,aes(x=JobRole,fill=factor(JobSatisfaction)))+
  geom_bar(width=0.5)+
  scale_fill_manual(values=blues)

#How people in our company are satisfied about their job by Department
greens <- brewer.pal(4, "Greens")

ggplot(df,aes(x=Department,fill=factor(JobSatisfaction)))+
  geom_bar(position="dodge",aplha=0.8)+
  scale_fill_manual(values=greens)

#Performance
ggplot(df,aes(x=Department,fill=OverTime))+
  geom_bar(width = 0.3) # The overtime rate has been controlled in a positive way.

  
```
## Q2 Which department has the highest attrition rate
```{r}
df$Attrition<-ifelse(df$Attrition=="Yes",1,0)

df_attrition_dept<-df%>%
  group_by(Department)%>%
  summarise(Attritionrate=mean(Attrition))
df_attrition_dept

ggplot(df_attrition_dept,aes(x=Department,y=Attritionrate,fill=Department))+
  geom_col(width = 0.5)
# Sales has the highest Attrition rate
```
## Q3 What’s the relationship between HourlyRate and MonthlyIncome ?
## New Q3 What's the relationship between the variables relevant to pay?;
# They are :
1. "DailyRate"
2. "HourlyRate"
3. "MonthlyIncome"
4. "MonthlyRate"
5. "PercentSalaryHike"
6. "TotalWorkingYears"
7. "YearsAtCompany"   
```{r}
library(corrplot)
df_payrate<-df[,c(4,13,19,20,24,29,32)]
COR<-cor(df_payrate)

pval <- psych::corr.test(df_payrate, adjust="none")$p

#corrplot for correlation coefficients
corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="lower",        
tl.col="blue", tl.cex=0.8, tl.srt=20, 
        addCoef.col="black", addCoefasPercent = TRUE,
        )
#corrplot for statistical significance
corrplot(COR, order="AOE",method="ellipse",type="lower", p.mat=pval, insig="p-value", 
         tl.pos="lt", sig.level=0.05,pch=20,pch.col="red",tl.cex=0.8, tl.srt=20)


#3 of the 21 correlation coefficients are statistically significant. totalworkingyears,monthlyincome and yearsatcompany are moderately to strongly correlated. This indicates thar the more time you stayed at IBM, more working experience you have, the more monthly income you can have.
```

#
```{r}

```


```{r}
library(ggplot2)
ggplot(df,aes(x=Age,fill=Gender))+
  geom_bar()

#dummy coding: if attrition is "Yes", coded as 0, whereas "No" as 1
df$Attrition<-ifelse(df$Attrition=="No",1,0)

#Attrition Rate
1-mean(df$Attrition) # 16.1% attrition seems okay, 10% is optimal.

#Check whether it is a good attrition or bad
df_attrition<-filter(df,df$Attrition==0)

t<-table(df_attrition$PerformanceRating,df_attrition$Department)
class(t)

ggplot(t,aes())+
  coord_polar("y", start=0)

prop.table(x=t,margin = 1)


#df%>%
#  group_by(Department)%>%
#  summarise(Attrition_rate=sum(df$Attrition))


```



```{r}
nrow(df)*0.75
sample_size<-sample(nrow(df),nrow(df)*0.75)

#set up training and 
training<-df[sample_size,]
testing<-df[-sample_size,]
pre_prune<-rpart.control(maxdepth = 30,cp=0)
model<-rpart(Attrition~.,training,method="class",control = pre_prune)



#post_pruning to get a more accurate model
plotcp(model)
post_prune<-prune(model,cp=0.013)

#draw a plot of 
rpart.plot(post_prune)
printcp(post_prune)


# test the model accuracy
testing$pre<-predict(post_prune,testing,"class")
mean(testing$pre==testing$Attrition)




```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
